Refinement for open automata

### 2021-05-05 2021-Refinement/report.tex
More comments

In the state \(RC\) the composition makes 2 transition \Ludo{from}{be more precise} the \(\tau\) transition but both have a false guard; let's look at one of them.
\nmm{\OT{RC}{\mpar{R, 3, C}}{\tau}{\mbrc{}}{\top \wedge \top \wedge c < t \wedge \act{set}\mpar{x} = \act{tick} \wedge \theta\mpar{x} = \delta\mpar{x_C}}{\mbrc{c \gets c + 1}}} is made with \nmm{\OT{2}{3}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller and \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
Obviously the guard is false so  it is equivalent to have no transition.
Also the guard  can be simplified.
In this article the guard and transition are always simplified except for the sake of illustrating a concept that is lost in the simplification as it is the case here.
We can see that the \(x\) variables from the specification, the controller and the register are in conflict.
To resolve the conflict we renamed the variables.

In the state \(RS\) the composition makes 1 transition from the \(\tau\) transition.
This transition is described before but let's look at the details of the composition.
The obtained guard is \(\top \wedge \top \wedge \top \wedge \theta\mpar{x} = \theta\mpar{17} \wedge \act{set}\mpar{x} = \act{set}\mpar{x_R}\) and the obtained new values of the variables is \(\mbrc{t \gets x_R, c \gets 0}\).
We can see here how value passing work: the only possible value of \(x\) and \(x_R\) making the guard hold is \(17\), this value has been synchronised from the controller to \(x\) then to the second occurence of \(x\) to the distinct \(x\) of the register.
This guard and value update can be simplified to the one shown on the drawing.
Note that \nmm{\OT{S}{C}{\act{set}\mpar{t}}{\mbrc{}}{\top}{\mbrc{c \gets 0}}} wouldn't have worked as a transition for the controller because \(t\) is fixed before the transition and its value cannot be set outside of the variable update, hence the use of a local variable \(x\).
\end{exi}

\section{What is a refinement for Open Automata}\label{sec:proofelts}
There are several properties that we may want from a refinement relation.
Depending on these properties, several kind of refinement may be used.
For example if we are interested in being able to produce the same sequence of action we may want to use trace set inclusion as a refinement.
Here the main properties that we want are related to composition and action refinement.
Because of composition, a relation as strong as simulation refinement must be used as observed in the articles about FH-bisimulation \cite{henrio:01055091}.

@@

The important properties we will consider here are:
\begin{defi} A relation \(\leq\) is
\defitem \textbf{reflexive} iff \(\forall a, a \leq a\);
\defitem \textbf{transitive} iff \(\forall a\, b\, c, a \leq b \wedge b \leq c \implies a \leq c\);
\defitem \textbf{a preorder} iff it is reflexive and transitive;
\defitem \textbf{complete w.r.t.\@ composition} iff \(\forall a\, b, a\mbrk{b} \leq a\);
\defitem \textbf{correct w.r.t.\@ composition} iff \(\forall a\, b, a \leq b \implies \exists c, a \overset{FH}= b\mbrk{c}\);
\defitem \textbf{context equivalent for composition} iff \(\forall a\, b\, c, a \leq b \implies a\mbrk{c} \leq b\mbrk{c}\);
\defitem \textbf{congruent for composition} iff \(\forall a\, b\, c, a \leq b \implies c\mbrk{a} \leq c\mbrk{b}\);
\defitem \textbf{compatible with composition} iff \(\forall a\, b\, c\, d, a \leq b \wedge c \leq d \implies c\mbrk{a} \leq d\mbrk{b}\);
\defitem \textbf{compatibile with FH-bisimulation} iff \(\forall a\, b\, c, d, a \overset{FH}= b \wedge c \overset{FH}= d \wedge a \leq c \implies b \leq d\).
\end{defi}
Reflexivity, transitivity and preorder are classical properties on relations.
Completeness w.r.t.\@ composition is the fact that every composition is considered a refinement.
Correctness w.r.t.\@ composition is the fact that a refinement correspond to the left automaton being equivalent to some composition of the right automaton.
Context equivalence is the refinement being compatible with composition with the same automaton.
Congruence is the refinement being compatible with being composed in the same automaton.
Compatibility with composition is the two latter at the same time.
Finally compatibility with FH-bisimulation is the refinement not being able to distinguish two FH-bisimilar automata.

>>>>

In the state \(RC\) the composition makes 2 transition \Ludo{from}{be more precise} the \(\tau\) transition but both have a false guard; \Ludo{Let's look}{avoid informal english} at one of them.
\nmm{\OT{RC}{\mpar{R, 3, C}}{\tau}{\mbrc{}}{\top \wedge \top \wedge c < t \wedge \act{set}\mpar{x} = \act{tick} \wedge \theta\mpar{x} = \delta\mpar{x_C}}{\mbrc{c \gets c + 1}}} is made with \nmm{\OT{2}{3}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller and \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
Obviously the guard is false so  it is equivalent to have no transition.
\Ludo{rest of the paragraph}{why so much discussion on a useless transition? did I miss something?} Also the guard  can be simplified.
In this article the guard and transition are always simplified except for the sake of illustrating a concept that is lost in the simplification as it is the case here.
We can see that the \(x\) variables from the specification, the controller and the register are in conflict.
To resolve the conflict we renamed the variables.

In the state \(RS\) the composition makes 1 transition from the \(\tau\) transition.
This transition is described before but \Ludo{Let's look}{avoid informal english} at the details of the composition.
The obtained guard is \(\top \wedge \top \wedge \top \wedge \theta\mpar{x} = \theta\mpar{17} \wedge \act{set}\mpar{x} = \act{set}\mpar{x_R}\) and the obtained new values of the variables is \(\mbrc{t \gets x_R, c \gets 0}\).
We can see here how value passing works: the only possible value of \(x\) and \(x_R\) making the guard hold is \(17\), this value has been synchronised from the controller to \(x\) then to the second occurrence of \(x\) to the distinct \(x\) of the register.
This guard and \Ludo{value update}{what do you mean?} can be simplified to the one shown on the drawing.
Note that \nmm{\OT{S}{C}{\act{set}\mpar{t}}{\mbrc{}}{\top}{\mbrc{c \gets 0}}} wouldn't have worked as a transition for the controller because \(t\) is fixed before the transition and its value cannot be set outside of the variable update, hence the use of a local variable \(x\).
\end{exi}

\section{What is a refinement for Open Automata}\label{sec:proofelts}
There are several properties that we may want from a refinement relation.
Depending on these properties, several kinds of refinement may be used.
For example if we are interested in producing the same sequence of actions we may want to use trace set inclusion as a refinement.
Here the main properties that we want are related to composition and action refinement.
Because of composition, a relation as strong as simulation refinement must be used as observed in the articles about FH-bisimulation \cite{henrio:01055091}.

@@

The important properties we will consider here are:
\begin{defi} A relation \(\leq\) is
\defitem \textbf{reflexive} iff \(\forall a, a \leq a\);
\defitem \textbf{transitive} iff \(\forall a\, b\, c, a \leq b \wedge b \leq c \implies a \leq c\);
\defitem \textbf{a preorder} iff it is reflexive and transitive;
\defitem \textbf{\Ludo{complete}{not sure whyy this is called completeness} w.r.t.\@ composition} iff \(\forall a\, b, a\mbrk{b} \leq a\);
\defitem \textbf{\Ludo{correct}{This eems very difficult to achieve so calling it correctness and implying that the other relations are not interesting is very strong} w.r.t.\@ composition} iff \(\forall a\, b, a \leq b \implies \exists c, a \overset{FH}= b\mbrk{c}\);
\defitem \textbf{context \Ludo{equivalent}{equivalent is strange here as there is no equivalence as far as I can see} for composition} iff \(\forall a\, b\, c, a \leq b \implies a\mbrk{c} \leq b\mbrk{c}\);
\defitem \textbf{congruent for composition} iff \(\forall a\, b\, c, a \leq b \implies c\mbrk{a} \leq c\mbrk{b}\);
\defitem \textbf{compatible with composition} iff \(\forall a\, b\, c\, d, a \leq b \wedge c \leq d \implies c\mbrk{a} \leq d\mbrk{b}\);
\defitem \textbf{compatibile with FH-bisimulation} iff \(\forall a\, b\, c, d, a \overset{FH}= b \wedge c \overset{FH}= d \wedge a \leq c \implies b \leq d\).
\end{defi}
Reflexivity, transitivity and preorder are classical properties on relations.
Completeness w.r.t.\@ composition is the fact that every composition is considered a refinement.
Correctness w.r.t.\@ composition is the fact that a refinement correspond to the left automaton being equivalent to some composition of the right automaton.
Context equivalence is the refinement being compatible with composition with the same automaton.
Congruence is the refinement being compatible with being composed in the same automaton.
Compatibility with composition is the two latter at the same time \Ludo{?}{(plus transitivity?) it depends what you mean by ``at the same time''}.
Finally compatibility with FH-bisimulation is the refinement not being able to distinguish two FH-bisimilar automata.
